The Types and Driving Forces of Modifications in the Evolution of an API

### 2015-05-13 thesis/paper.tex
Minor changes to RQ1 sections

All of the identified changes have been given a type based the nature of the change according to the Types of Software Evolution and Software Maintenance decision tree \cite{chapin2001types}. In total, thirteen unique changes were found. These changes are shown in Table \ref{table:categories}.

@@

\smallskip \noindent
\textbf{Corrective  }
%Define what kinds of changes these are
Corrective changes are identified as types of change that fix broken functionality.
%Data and explanation
No corrective changes were identified between V1.0 and V2.0. There are two main reasons that we did not find any changes of this type. Firstly, the majority of changes were additions or re-implementations of functionality. This meant that most of the existing code was either removed, replaced or deprecated resulting in little to no obvious fixes. Secondly, we only had access to the public facing API and not the implementation itself. We therefore could not inspect the inner workings of the functionality that has remained between versions to determine if any corrective change took place.
%Conclusion
We therefore conclude that the evolution of the API was not driven by corrective changes.
\smallskip \noindent
\textbf{Reductive  }
%Definition
Reductive changes are defined as changes that remove or reduce functionality. This is distinctive from a deprecation as reductive types of change remove functionality completely.
%Data & Explanation
Only a single reductive type of change was identified. This reductive change removed functionality that provided the ability to buffer individual images for closer analysis by a user. Given that it is more common to refactor APIs and not remove functionality \cite{dig2005role, xing2006refactoring} this was a surprising result. 
%Don't know if I should discuss the finding in this much depth here?
%If OK - Talk more about the business side of things, and how we couldn't identify this change ourselves
It was revealed during the interview that the functionality was removed because the case company no longer wished to support this feature, due to hardware constraints. 
%Conclusion
As this was the only reductive change found, we conclude that the evolution of the API was not primarily driven by reductive changes.

@@

\noindent
\textbf{Functionality is a prime driver of API evolution  }  %Reviewed by JW
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found that a vast majority of the changes either added functionality or were a direct consequence of increased functionality. The main driver for these changes was the pressure from API users, but pressure from the case company developers who utilise the API also contributed. This focus upon functionality was also emphasised partially due to a business goal. The more features the API can provide, the more applications developers can create for the cameras. This has a domino effect of increasing the appeal and therefore value of the products.

\smallskip \noindent 
\textbf{Maintainability does not drive change  }  %Reviewed by JW
While maintainability is important, it is not a primary driving factor for change. We found that where changes that effected maintainability, functionality had also been added or modified. Of the changes identified, none were considered to be primarily groomative. Groomative motives were only found to be secondary, and were never the main motive for a change. Changes to the example code are an instance of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

@@

\noindent
\textbf{A majority of the laws apply  } %Reviewed by JW
Sufficient evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

>>>>

All of the identified changes have been given a type based the nature of the change according to the Types of Software Evolution and Software Maintenance decision tree \cite{chapin2001types}. In total, thirteen unique changes were found.

@@

\smallskip \noindent
\textbf{Corrective  }
%Define what kinds of changes these are
Corrective changes are identified as types of change that fix broken functionality.
%Data and explanation
No corrective changes were identified between v1.0 and v2.0. There are two main reasons that we did not find any changes of this type. Firstly, the majority of changes were additions or re-implementations of functionality. This meant that most of the existing code was either removed, replaced or deprecated, resulting in little to no obvious fixes. Secondly, we only had access to the public facing API and not the implementation itself. We therefore could not inspect the inner workings of the functionality that has remained between versions to determine if any corrective change took place.
%Conclusion
We therefore conclude that the evolution of the API was not driven by corrective changes.
\smallskip \noindent
\textbf{Reductive  }
%Definition
Reductive changes are defined as changes that remove or reduce functionality. This is distinctive from a deprecation as reductive types of change remove functionality completely.
%Data & Explanation
Only a single reductive type of change was identified. This reductive change removed functionality that provided the ability to buffer individual images for closer analysis by a user. Given that it is more common to refactor APIs and not remove functionality \cite{dig2005role, xing2006refactoring} this was a surprising result. 
%Don't know if I should discuss the finding in this much depth here?
%If OK - Talk more about the business side of things, and how we couldn't identify this change ourselves
It was revealed during the interview that the functionality was removed because the case company no longer wished to support this feature, due to hardware constraints. 
%Conclusion
As this was the only reductive change found, we conclude that the evolution of the API was not primarily driven by reductive changes.

@@

\noindent
\textbf{Functionality is a prime driver of API evolution  }  %Reviewed by JW
Overwhelmingly, the primary driving factor behind API evolution was the desire for new API functionality. We found that a vast majority of the changes either added functionality or were a direct consequence of increased functionality. The main driver for these changes was the pressure from API users, but pressure from the case company developers who utilise the API also contributed. This focus upon functionality was also emphasised partially due to a business goal. The more features the API provides allows for more applications to be created for the cameras. This has a domino effect of increasing the appeal and value of the products the case company produces.

\smallskip \noindent 
\textbf{Maintainability does not drive change  }  %Reviewed by JW
While maintainability is important, it is not a primary driving factor for change. We found that where changes that effected maintainability, functionality had also been added or modified. Of the changes identified, none were considered to be primarily groomative. Groomative motives were only found to be secondary, and were never the main motive for a change. Changes to the example code are an example of this. The example code was updated purely to reflect the changes in functionality and not as a groomative effort.

@@

\noindent
\textbf{A majority of the laws apply  } %Reviewed by JW
Evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.
